---
title: 'Homework 3: Fantastic distributions and where to find their PDFs'
author: "Yen-Chung Chen"
date: "2022-09-28"
output: html_document
---
### DUE: 10/04/22

### Please submit all your functions and answers in one file.

#### Q1. Let X be a discrete random variable whose probability mass function is described in the table given here.

|||||
|----|---|---|---|
| x  | -1| 0 | 1 |
|f(x)|1/8|6/8|1/8|

Please find:

1. P(X = 0) (2%)

```{r}

```

2. P(X < 10) (2%)

```{r}

```

3. P(X >= 0) (2%)

```{r}

```

4. Define a function `dcustom()` in R that represents the probability mass
function defined by the table (4%).

```{r}

```

#### Q2. You are working on a mark-recapture study of boreal toads, Bufo boreas. You predict that there is a 40% chance of capturing a marked toad for each of 30 traps that you establish.

1. What is the probability that exactly 20 marked toads will be found in the 30
traps? Calculate this by hand, and confirm your calculation using R. (5%)

```{r}

```

2. What is the probability that up to 20 marked toads will be found? Use R. (5%)

```{r}

```

#### Q3. A herd of mountain goats is believed to contain 30 males and 20 females. If, for a single capture it is equally easy to capture either male or female goats, what is the probability of capturing less than or equal to 3 females in 10 captures if you don’t release them between captures? (5%)

> Use R to calculate and explain your rationale.


```{r}

```

#### Q4. Assume the dog breeder for NYPD has 300 beagles and 700 golden retrievers, and 180 beagles and 20 golden retrievers were selected to train as sniffer dogs.

1. If the selection of sniffer dogs is random, what is the probability to have
more than or equal to 180 beagles to be selected? (5%)

```{r}

```

2. Assuming the selection process is random, simulate the process of selecting
200 sniffer dogs from the breeder for 100 times in R. What is the average number
of beagles you select? (5%)

Keep this line in the answer chunk together with your answer: `set.seed(1234)`.
`set.seed()` set the *seed* for randomization in R, so if you have `set.seed()`
run together with an otherwise random process, the result will be consistent,
which makes grading easier.

```{r}
set.seed(1234)

# Assign your simulation to an object before calculating the mean because you
# will need it in the next question.

```

3. With the simulation above in (2), how many times does your simulation
select > 60 beagles? (5%)

```{r}

```

4. Is the result of the simulation similar to what you get if you address the
question with `phyper()`? Scale up the simulation to 10,000 repeats. Does the
number of repeated simulation changes how consistent the simulation is to
`phyper()`? Describe how you assess similarity. (5%)

```{r}

```

#### Q5. Stanford–Binet IQ test scores are normally distributed with a mean μ, of 100, and a standard deviation σ, of 16. Let Stanford–Binet IQ be described by the random variable X. Use `pnorm()` and `qnorm()` to find the answers.

1. P(X = 50). Justify your answer. (5%)

```{r}

```

2. P(X > 100). Justify your answer. (5%)

```{r}

```

3. 20% of test scores are less than or equal to what IQ? (5%)

```{r}

```

4. 70% of test scores are more than what IQ? (5%)

```{r}

```

#### Q6. It is believed that an average honeycrisp apple weighs 136g. Assuming the weight of honeycrisp apples are normally distributed with a mean of 136g and a standard deviation of 10g, and that you collected 8 apples from an orchard that weighs 901g together

1. What is the probability that apples that are lighter than or as heavy as the
ones you are collecting have the same weight distribution as honeycrisp apples
do? (5%)

```{r}

```


2. Assuming you have an orchard of normally distributed honeycrisp apples with
mean weight = 136g and sd = 10g, how heavy is an apple that I have to bring so
it will be heavier than 90% of the apples in your orchard? (5%)

```{r}

```


#### Q7. RNA-seq experiments can be roughly seen as a Poisson process of RNA transcripts being detected. Assuming that there are 5 transcripts for RPL23, a gene encoding a 60S subunit, per 1 million transcripts in a cell type that you sequenced.

1. If you sequence 6 million transcripts, what is the probability that you see
more than 40 RPL23 transcripts? (5%)

```{r}

```

2. Assuming that transcripts are detected as a Poisson process, with a
2/1,000,000 (2 in 1 million) detection rate, simulate the process of sequencing
1 million transcripts for 100 times in R. What is the mean and variance
(with `var()`) of the number of transcripts detected? (5%)

Keep this line in the answer chunk together with your answer: `set.seed(1234)`.
`set.seed()` set the *seed* for randomization in R, so if you have `set.seed()`
run together with an otherwise random process, the result will be consistent,
which makes grading easier.

```{r}
set.seed(1234)
```

3. Alternatively, if we consider the detection of the transcript as a Bernoulli
process with a probability of 2/1,000,000, we can simulate (2) as a binomial
process with a size of 1 million. Repeat this simulation 100 times, and
calculate the mean and variance. Describe how the result compares with (2). (5%)

Keep this line in the answer chunk together with your answer: `set.seed(1234)`.
`set.seed()` set the *seed* for randomization in R, so if you have `set.seed()`
run together with an otherwise random process, the result will be consistent,
which makes grading easier.

```{r}
set.seed(1234)
```

4. Scale your simulation up to 10,000 times, and plot the transcript number
detected as a histogram with 20 bins. (5%)

#### Q8. Assume that the number of sightings for bald eagles (Haliaeetus leucocephalus) at a site over a 4-week period is Poisson distributed (temporally random) with a mean λ = 10. What is the probability of having 15 sightings over a 4-week period?

(5%)

```{r}

```

